---
title: "Crop suitability analysis"
author: "Dr. Burchfield"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: flatly
    toc: yes
    toc_float: true
    code_folding: hide
---

This doc contains the scripts to run the biophysical and full (biophysical + census) RF analyses. It also contains code used to assess the relative performance of different variable sets in predicting held-out data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=F, warning=F)
```

```{r message=F, warning=F}
set.seed(101)
source("CS_load.R")
source("CS_func.R")
select <- dplyr::select

# prep all the data for Census and bioclim runs
full_list <- list()
train_list <- list()
test_list <- list()
geom_list <- list()

for (c in 1:length(crop_names)) {

  crop_name <- crop_names[[c]]
  
  full <- apmc %>% 
    filter(CROP == crop_name) %>%
    select(AP, IRR, varlist[["Full"]]) %>%
    drop_na() %>%
    mutate(AP = as.factor(AP))
 
  geom_list[[crop_name]] <- st_geometry(full)
  st_geometry(full) <- NULL
  full_list[[crop_name]] <- full

  set.seed(1)
    
  # hold out 25% of the data, randomly selected in space and time:
  random_rn <- sample(nrow(full), ceiling(nrow(full)*.25))
  train_list[[crop_name]] <- full[-random_rn,]
  test_list[[crop_name]] <- full[random_rn,]
  
}
```


# Performance 

```{r eval=F}
# This code generates error distributions by running on multiple test/training subsets of data:
# and run the models for multiple iterations
library(foreach)
library(doParallel)

num_cores <- detectCores() - 6
clus <- makeCluster(num_cores)
registerDoParallel(clus)
run_those_models(crop_names, nint = 10)
stopCluster(clus)
```


# Historical random forests models 

Prepare data (built in `CS_clean.R` and described in `CS_data_construction.html`). Run on the subset of biophysical parameters (climate + soil + topography) and on the set of "all predictors" that includes inputs, farm resources, and farmer characteristics.


## Climate only

```{r}
out_ext <- "_BIO.RDS"
ntrain_list <- map(train_list, ~select(., AP, varlist[["Biophysical"]]))
```

Run `randomForest()`:

```{r eval=F}
run_historical_models(mtry_dir = "./out/mtry.RDS", out_ext = out_ext, data_list = ntrain_list)
```

Performance on testing data across models:

```{r}
perf <- rf_performance(test_list = test_list, out_ext = out_ext)
DT::datatable(perf)
```

Variable importance:

```{r echo=F}
imp_final <- build_imp(out_ext)

ggplot(imp_final %>% filter(CROP != "SWHEAT"), aes(x=reorder(varnames, MeanDecreaseGini), y=MeanDecreaseGini)) + 
    geom_point() +
    geom_segment(aes(x=varnames,xend=varnames,y=0,yend=MeanDecreaseGini)) +
    ylab("Mean decrease Gini") +
    xlab("") +
    coord_flip() +
    project_theme +
    facet_wrap(~ CROP, ncol = 3, nrow=3) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "bottom")
```


## All predictors

```{r}
out_ext <- "_FULL.RDS"
ntrain_list <- map(train_list, ~select(., AP, varlist[["Full"]]))
```

Run `randomForest()`:

```{r eval=F}
run_historical_models(mtry_dir = "./out/mtry_FULL.RDS", out_ext = out_ext, data_list = ntrain_list)
```

Performance on testing data across models:

```{r}
perf <- rf_performance(test_list = test_list, out_ext = out_ext)
DT::datatable(perf)
```

Variable importance:

```{r echo=F}
imp_final <- build_imp(out_ext)

ggplot(imp_final %>% filter(CROP != "SWHEAT"), aes(x=reorder(varnames, MeanDecreaseGini), y=MeanDecreaseGini)) + 
    geom_point() +
    geom_segment(aes(x=varnames,xend=varnames,y=0,yend=MeanDecreaseGini)) +
    ylab("Mean decrease Gini") +
    xlab("") +
    coord_flip() +
    project_theme +
    facet_wrap(~ CROP, ncol = 3, nrow=3) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "bottom")

# ggsave(paste0("./figs/by_crop_vip_", substr(out_ext, 1, 3), ".png", plot = last_plot(), width = 12, height = 8) # fix varnames
```

# Historical suitability prediction

## Biophysical

```{r eval=F}
out_ext <- "_BIO.RDS"
ntrain_list <- map(train_list, ~select(., AP, varlist[["Biophysical"]]))
build_historical_grids(out_ext, vars = colnames(ntrain_list[[1]])) # takes longer with hires data
```

```{r}
out_ext <- "_BIO.RDS"
hv_clim <- build_hist_vis(out_ext)
hv_clim[[1]]
```

## All predictors

```{r}
out_ext <- "_FULL.RDS"
ntrain_list <- map(train_list, ~select(., AP, varlist[["Full"]]))
build_historical_grids(out_ext, vars = colnames(ntrain_list[[1]]))
```

```{r}
out_ext <- "_FULL.RDS"
hv_full <- build_hist_vis(out_ext)
hv_full[[1]]
```

## Delta suitability

```{r}
all_ras <- hv_full[[2]] - hv_clim[[2]]

names(all_ras) <- clean_crop_names[!(clean_crop_names %in% "Spring wheat")]
  
state_sp <- as(states, "Spatial")
state_sp <- spTransform(state_sp, all_ras@crs)
  
coul2 <- colorRampPalette(brewer.pal(11, "BrBG"))
  
  
diff <- levelplot(all_ras, 
                         margin = F,
                         xlab = "",
                         ylab = "All predictors - climate and topography",
                colorkey=list(space="bottom"),
                         # xlab = list(names(all_ras), space = "top"),  # winter wheat
                         # ylab=list(c("2081-2100", "2061-2080", "2041-2060", "2000-2020"), rot=0), 
                         scales = list(draw=F), 
                         layout = c(3, 2), 
                         names.attr = names(all_ras), #rep("", dim(allvis)[3]),
                         col.regions = coul2,
                  at = seq(-0.8, 0.8, length.out = 11)) + # inferno viridis
    latticeExtra::layer(sp.polygons(state_sp, lwd = 0.7, col = "#696969"))
  

diff
```

## Visualizing historical error

I built a set of 2.5 million randomly allocated points. I extracted the actual CDL pixel values through time to each pixel and flagged whether each crop appeared from 2008 to 2019: 

```{r eval=F}
# sample points
rpt <- st_as_sf(spsample(as(roi_rpj, "Spatial"), n = 2500000, "random")) # big n needed to ensure 10K AP for each crop
saveRDS(rpt, "./data/sample_points.RDS")

# extract CDL designated crop to points through time
years <- 2008:2019
rpt <- readRDS("C:/Users/eburchf/Desktop/Data/CDL/CS_project/sample_points.RDS")
cdl_list <- Sys.glob("C:/Users/eburchf/Desktop/Data/CDL/*.img")
for (y in 1:length(cdl_list)) {

  print(years[[y]])
  cdl <- raster(cdl_list[[y]])
  rpts <- rpt

  out <- raster::extract(cdl, rpts, df=T)
  saveRDS(out, paste0("C:/Users/eburchf/Desktop/Data/CDL/CS_project/year", years[[y]], ".RDS"))

}

cdl_extract <- cbind.data.frame(lapply(Sys.glob("C:/Users/eburchf/Desktop/Data/CDL/CS_project/year*.RDS"), FUN = readRDS)) %>% select(-c(ID))
saveRDS(cdl_extract, "./out/cdl_out.RDS")

# Detect crops in CDL data
cdl_extract <- readRDS("./out/cdl_out.RDS")
cdle <- data.frame()[1:nrow(cdl_extract), ]
cdle$CORN <- apply(cdl_extract,1,function(x) {sum(x %in% crop_dc_list[[1]],na.rm = TRUE)})
cdle$COTTON <- apply(cdl_extract,1,function(x) {sum(x %in% crop_dc_list[[2]],na.rm = TRUE)})
cdle$SOY <- apply(cdl_extract,1,function(x) {sum(x %in% crop_dc_list[[3]],na.rm = TRUE)})
cdle$SWHEAT <- apply(cdl_extract,1,function(x) {sum(x %in% crop_dc_list[[4]],na.rm = TRUE)})
cdle$WWHEAT <- apply(cdl_extract,1,function(x) {sum(x %in% crop_dc_list[[5]],na.rm = TRUE)})
cdle$ALFALFA <- apply(cdl_extract,1,function(x) {sum(x %in% crop_dc_list[[6]],na.rm = TRUE)})
cdle$HAY <- apply(cdl_extract,1,function(x) {sum(x %in% crop_dc_list[[7]],na.rm = TRUE)})

# Count if they are more than threshold, save as sf
tha <- 0 
actual <- cdle %>%
  mutate(CORN_ACT = ifelse(CORN > tha, "CORN", "OTHER"),
         ALFALFA_ACT = ifelse(ALFALFA > tha, "ALFALFA", "OTHER"),
         COTTON_ACT = ifelse(COTTON > tha, "COTTON", "OTHER"),
         HAY_ACT = ifelse(HAY > tha, "HAY", "OTHER"),
         SOY_ACT = ifelse(SOY > tha, "SOY", "OTHER"),
         SWHEAT_ACT = ifelse(SWHEAT > tha, "SWHEAT", "OTHER"),
         WWHEAT_ACT = ifelse(WWHEAT > tha, "WWHEAT", "OTHER"))
st_geometry(actual) <- st_geometry(rpt)
actual <- st_transform(actual, hs@crs)
saveRDS(actual, "./out/cdl_extract.RDS")
```

Then I rasterized the CDL data and compared it to the predicted surfaces. Here, the threshold is 0.5:

```{r eval=F}
# build categorical error rasters
bio_th <- find_threshold("_BIO.RDS", index = "Accuracy")
build_error_rasters("BIO", source = "_DayMet", fact = 1, thdf = bio_th)

full_th <- find_threshold("_FULL.RDS", index = "Accuracy")
build_error_rasters("FULL", source = "_DayMet", fact = 1, thdf = full_th) #0.5
```

Build rasterized depictions of error by coarsening rasters to categories:

```{r eval=F}
clim_ras <- stack(Sys.glob("./out/hist_error/cat*_BIO_1_CSTH.tif")) 
names(clim_ras) <- c("Alfalfa", "Corn", "Cotton", "Hay", "Soy", "Wheat")

full_ras <- stack(Sys.glob("./out/hist_error/cat*_FULL_1_CSTH.tif"))
names(full_ras) <- c("Alfalfa", "Corn", "Cotton", "Hay", "Soy", "Wheat")

cols <- c("#228B22", "red", "blue", "light gray")
state_sp <- spTransform(state_sp, clim_ras@crs)

sras <- stack()

for (i in 1:dim(clim_ras)[[3]]) {
  
  r <- clim_ras[[i]]
  r <- ratify(r)
  rat <- data.frame(ID = c(1, 2, 3, 4))
  rat$TYPE <- NA
  rat$TYPE[rat$ID == 1] <- "Correctly classified suitability" # green
  rat$TYPE[rat$ID == 2] <- "Falsely classified suitability" # red
  rat$TYPE[rat$ID == 3] <- "Falsely classified unsuitability" # blue
  rat$TYPE[rat$ID == 4] <- "Correctly classified unsuitability" # gray
    
  levels(r) <- rat
  sras <- stack(sras, r)
  
}
  
clim_viz <- levelplot(sras, 
                      margin = F,
                      xlab = "",
                      ylab = "Biophysical",
                      names.attr = names(clim_ras),
                      colorkey=F,
                      layout = c(3,2),
                      #  ylab=list(c("SSP585", "SSP245"), space = "top"),  # winter wheat
                      # xlab=list(c("2000-2020", "2041-2060", "2061-2080", "2081-2100"), rot=0),
                      scales = list(draw=F), 
                      #  layout = c(4, 2), 
                      # names.attr = rep("", dim(allvis)[3]),
                      col.regions = cols) + # inferno viridis
      layer(sp.polygons(state_sp, lwd = 0.7, col = "#696969"))


g1 <- grid.arrange(clim_viz, ncol=1, nrow=1)
ggsave("./figs/cat_error_BIO_1_CSTH.png", g1, width=8, height=8)


sras <- stack()

for (i in 1:dim(clim_ras)[[3]]) {
  
  r <- full_ras[[i]]
  r <- ratify(r)
  rat <- data.frame(ID = c(1, 2, 3, 4))
  rat$TYPE <- NA
  rat$TYPE[rat$ID == 1] <- "Correctly classified suitability" # green
  rat$TYPE[rat$ID == 2] <- "Falsely classified suitability" # red
  rat$TYPE[rat$ID == 3] <- "Falsely classified unsuitability" # blue
  rat$TYPE[rat$ID == 4] <- "Correctly classified unsuitability" # gray
    
  levels(r) <- rat
  sras <- stack(sras, r)
  
}

full_viz <- levelplot(sras, 
                      margin = F,
                      xlab = "",
                      ylab = "All predictors",
                      names.attr = names(full_ras),
                      layout = c(3,2),
                      colorkey=F,
                      #  ylab=list(c("SSP585", "SSP245"), space = "top"),  # winter wheat
                      # xlab=list(c("2000-2020", "2041-2060", "2061-2080", "2081-2100"), rot=0),
                      scales = list(draw=F), 
                      #  layout = c(4, 2), 
                      # names.attr = rep("", dim(allvis)[3]),
                      col.regions = cols) + # inferno viridis
      layer(sp.polygons(state_sp, lwd = 0.7, col = "#696969"))


g2 <- grid.arrange(full_viz, ncol=1, nrow=1)
ggsave("./figs/cat_error_FULL_1+CSTH.png", g2, width=8, height=8)

g3 <- grid.arrange(g1, g2, ncol = 1, nrow = 2)
ggsave("./figs/cat_error_1_CSTH.png", g3, width = 8, height = 10)
```


# Future suitability predictions

Build baseline reference historical suitability surface using attr_grids/5K_hist_attr merged with the ensemble biovars surfaces

For each scenario (then average suitability rasters):

```{r eval=F}
build_future_grids(out_ext = "_BIO.RDS", full_list = full_list)
```

Then compute mean suitability projection across models and sd:

```{r}
ssps <- c("245", "370", "585")
years <- c("2021-2040", "2041-2060", "2061-2080", "2081-2100")
crop_names_sub <- c("CORN", "COTTON", "SOY", "WWHEAT", "ALFALFA", "HAY")

for (i in 1:length(ssps)) {
  
  print(ssps[[i]])

  for (j in 1:length(crop_names_sub)) {
    
    print(crop_names_sub[[j]])
    
    y20 <- stack(Sys.glob(paste0("./out/future_grids/prob*ssp", ssps[[i]], "*", "2021-2040", "*", crop_names_sub[[j]], "*.tif")))
    y40 <- stack(Sys.glob(paste0("./out/future_grids/prob*ssp", ssps[[i]], "*", "2041-2060", "*", crop_names_sub[[j]], "*.tif")))
    y60 <- stack(Sys.glob(paste0("./out/future_grids/prob*ssp", ssps[[i]], "*", "2061-2080", "*", crop_names_sub[[j]], "*.tif")))
    y80 <- stack(Sys.glob(paste0("./out/future_grids/prob*ssp", ssps[[i]], "*", "2081-2100", "*", crop_names_sub[[j]], "*.tif")))
    
    y20m <- calc(y20, mean)
    y20s <- calc(y20, sd)
    
    y40m <- calc(y40, mean)
    y40s <- calc(y40, sd)
    
    y60m <- calc(y60, mean)
    y60s <- calc(y60, sd)
    
    y80m <- calc(y80, mean)
    y80s <- calc(y80, sd)
    
    ms <- stack(y20m, y40m, y60m, y80m)
    names(ms) <- c("2021-2040", "2041-2060", "2061-2080", "2081-2100")
    ss <- stack(y20s, y40s, y60s, y80s)
    names(ss) <- c("2021-2040", "2041-2060", "2061-2080", "2081-2100")
    
    writeRaster(ms, paste0("./out/future_ens/mean_stack_", ssps[[i]], "_", crop_names_sub[[j]], ".tif"), overwrite=T)
    writeRaster(ss, paste0("./out/future_ens/sd_stack_", ssps[[i]], "_", crop_names_sub[[j]], ".tif"), overwrite=T)

  }

}

```

